\documentclass[twocolumns]{IEEEtran}
\usepackage{tikz}
\tikzstyle{box}=[rectangle,draw=black, ultra thick, minimum size=1cm]
\usepackage{algorithm}
\usepackage{algpseudocode}

\author{Erdal Sidal Dogan, Alp Gokcek \\ MEF University \\ \today}
\title{Max-Subarray Sum Problem and Solution Algorithms}


\begin{document}
	\maketitle
	\begin{abstract}
		Maximum Subarray Sum is a well-known problem in the field of computer science. There are multiple number of solution algorithms with different complexities. In this paper, we demonstrated and compared 3 of these algorithms with quadratic, linear, and logarithmic complexities.
	\end{abstract}
\section{The Problem}
The Maximum Subarray Sum problem is the task of finding the contiguous subarray with largest sum in a given array of integers. Each number in the array could be positive, negative, or zero. For example: Given the array $[-2, 1, -3, 4, -1, 2, 1, -5, 4]$ the solution would be $[4, -1, 2, 1]$ with a sum of 6.
\section{Solutions}
\subsection{Brute-Force Approach}
This is the most intuitive solution to anyone. You basically traverse over the array and compare every possible combination of start and end index for the soultion array.\\ \\


\begin{tikzpicture}[transform canvas={scale=.5}, trim left=-1cm]
[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
box/.style={rectangle,draw=black, ultra thick, minimum size=0.5cm},
]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\foreach \x/\y in {0/9, 1/5,2/13,3/19,4/12,5/8,6/7,7/4,8/21,9/2,10/6,11/11}
\node[box] at (\x,0){\y};

\draw[->,very thick] (0,1.2) --  node[above,yshift=2mm]{start} (0,.7);
\draw[->,very thick] (1,1.2) --  node[above,yshift=2mm]{end} (1,.7);
\end{tikzpicture} 
\\ \\ \\
The \textit{start} index will be incremented by 1 everytime the \textit{end} index reaches the end of the array. Then the \textit{end} index will start from the element right next to the start element. At every iteration the sum between \textit{start} and \textit{end} indexes will be calculated. Hence, the maximum sum will be determined by computing every sum for the every possible sub-array. The complexity of this algorithm is $\mathcal{O}(n^3)$. With a little improvement we can convert this algorithm to be $\mathcal{O}(n^2)$. Instead of calculating the sum between two array indicies at every iteration from scratch, we know that the current sum will be the (current element + previous sum). Consequently, eliminating the loop which is used for calculating the sum from the algorithm will reduce the time complexity.
\newpage
\begin{algorithm}
	\caption{Brute-Force}
	\begin{algorithmic}[line]
		\State $n \leftarrow len(array)  $ 
		\State $max\_sum \leftarrow 0$ 
		\For {$i \leftarrow 1$ to $ n $} 
			\For{$j \leftarrow 1$ to $n$}
			\State $sum += array[j]$
			\If{$sum > max\_sum$}
				\State $max\_sum \leftarrow sum$ 
			\EndIf
			\EndFor
		\EndFor
	\end{algorithmic}
\end{algorithm}
\subsection{Divide \& Conquer Approach}
Another solution is to divide the array into half recursively and computing the max subarray sum for each half and the sub array for crossing both halfs. After calculating the summation for these 3 cases, we choose the largest one, thus we determine the maximum sub array.\\ 

\begin{tikzpicture}[transform canvas={scale=.5}, trim left=-2.5cm]
	[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	box/.style={rectangle,draw=black, ultra thick, minimum size=0.5cm},
	]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
	\foreach \x/\y in {0/9, 1/5,2/13,3/19,4/12,5/8,6/7,7/4}
	\node[box] at (\x,0){\y};
	
	%\draw[->,very thick] (0,1.2) --  node[above,yshift=2mm]{start} (0,.7);
	%\draw[->,very thick] (1,1.2) --  node[above,yshift=2mm]{end} (1,.7);
	\draw[->,very thick] (3.45,-.7) -- (1,-2);
	\draw[->,very thick] (3.55,-.7) -- (6,-2);
	
	\foreach \x/\y in {-1/9, 0/5,1/13,2/19}
	\node[box] at (\x,-3){\y};
	
	\foreach \x/\y in {5/12, 6/8,7/7,8/4}
	\node[box] at (\x,-3){\y};
	
	\draw[->,very thick] (0.45,-3.6) -- (-0.7,-4);
	\draw[->,very thick] (0.55,-3.6) -- (1.7,-4);
	
	\draw[->,very thick] (6.45,-3.6) -- (5.35,-4);
	\draw[->,very thick] (6.55,-3.6) -- (7.65,-4);
	
	\foreach \x/\y in {8/7,9/4}
	\node[box] at (\x,-5){\y};
	\foreach \x/\y in {4.5/12, 5.5/8}
	\node[box] at (\x,-5){\y};
	
	\foreach \x/\y in {-2/9,-1/5}
	\node[box] at (\x,-5){\y};
	\foreach \x/\y in {1.5/13,2.5/19}
	\node[box] at (\x,-5){\y};
\end{tikzpicture} 
\\ \\ \\ \\ \\ \\ \\  \\ 

\begin{algorithm}
	\caption{Divide \& Conquer}
	\begin{algorithmic}[line]
	\State \textbf{function} max\_crossing\_sub\_array$(array, l, m, h)$		
	\State$left\_max\_sum \leftarrow -100000$
	\State$sum\_l \leftarrow 0$
	\For{$ i \leftarrow m$ \textbf{downto} $l - 1$} 
		\State$sum\_l \leftarrow sum\_l + array[i]$
		\If{$sum\_l > left\_max\_sum$}
			\State $left\_max\_sum \leftarrow sum\_l$
		\EndIf
	\EndFor
	
	\State $sum\_r \leftarrow 0$
	\State $right\_max\_sum \leftarrow -100000$
	\For{$ j \leftarrow m + 1$ \textbf{to} $h + 1$} 
		\State$sum\_r \leftarrow sum\_r + array[j]$
		\If{$sum\_r > right\_max\_sum$}
			\State $right\_max\_sum \leftarrow sum\_r$
		\EndIf
	\EndFor
	\Return$left\_max\_sum + right\_max\_sum )$
	\end{algorithmic}
\end{algorithm}


\end{document}